---
title: "marker_coverage"
author: "Michael Overton"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: "hide"
---

***

## Per clone marker density


```{r}
SNPs_counts_Line <- LOH_SNPs %>%
  filter(Rep != "00") %>%
  count(Line, CHROM, POS, GT, name = "n_conv") %>%
  ungroup %>%
  mutate(per_clone = operate_by_factor_match(n_clones_LOH_Line[, c("Line", "n")], 
                                             data.frame(Line, n_conv),
                                             function(x, y) y/x)) 
m_cover_ID <- LOH_SNPs %>%
  filter(Rep != "00") %>%
  ungroup() %>%
  count(ID)

mean(m_cover_ID$n)
median(m_cover_ID$n)
range(m_cover_ID$n)

m_cover_ID %>% ggplot(aes(x = n)) + geom_histogram(binwidth = 500) 

```

## Marker density and LOH counts across MA lines

```{r}
mean_m_cover_Line <- SNPs_counts_Line %>% 
  group_by(Line, CHROM, POS) %>%
  summarize(n_m = sum(n_conv)) %>%
  ungroup() %>%
  # filter(n_m > 2) %>%
  count(Line)

mean_m_cover_Line <- merge(mean_m_cover_Line, LOH_Line, by = "Line")

mean_m_cover_Line %>% 
  ggplot() + geom_point(aes(x = n, y = tot_LOH)) +
  # xlim(0, NA) + 
  ylim(0, NA)



```



## Marker density and LOH counts across chromosomes

We want to know whether the differences in marker density are affecting 
our ability to detect LOH events and our LOH rate estimates. It appears that
marker density is lower in the middle of larger Chrs, especially IV. We can
ask what the marker density is per kilobase.

```{r}
m_cover_CHROM_Tx <- LOH_SNPs %>% 
  count(Tx_ID, CHROM) %>%
  mutate(per_bp = operate_by_factor_match(chrom_lengths_BY_df[, c("CHROM", "chrom_length")],
                                          data.frame(CHROM, n),
                                          function(x, y) y/x))
m_cover_CHROM_Tx$Tx_ID <- Recode_Tx_ID(m_cover_CHROM_Tx$Tx_ID, "Tx_ID")

m_cover_CHROM_Tx <- merge(m_cover_CHROM_Tx, 
                          LOHcounts_CHROM_mean[, c("Tx_ID", "CHROM", "mean_LOH")], 
                          by = c("Tx_ID", "CHROM"))

m_cover_CHROM_Tx %>% 
  ggplot() + geom_point(aes(x = per_bp, y = mean_LOH, color = Tx_ID)) +
  scale_color_manual(values = txPal)

m_cover_CHROM_Tx %>% 
  ggplot() + geom_point(aes(x = per_bp, y = mean_LOH, color = CHROM)) 

```


